Peer Assessment 1: Reproducible Research 
========================================================
Loading and processing the data
=================

```{r}
data<-read.csv( "activity.csv" , header=T, sep= ",")
```
Convert date to date class

```{r}
data$Date <- as.Date(data$date, "%Y-%m-%d")
```
Convert interval to a factor

```{r}
data$interval <- as.factor(data$interval)
```

Extract levels of 5-min intervals
```{r}
lev <- levels(data$interval)
```
Average total number of steps taken per day
```{r fig.width=9, fig.height=8}
library(plyr)
totalSteps <- tapply(data$steps, data$date, sum, na.rm=T)
avgSteps <- tapply(data$steps, data$date, mean, na.rm=T)
summary(totalSteps)
summary(avgSteps)
hist(totalSteps, breaks=10, col="green", main="Total Number of steps each day", xlab="Average Steps per day")
```
daily Average  activity pattern
```{r}
Steps = tapply(data$steps, data$interval, mean, na.rm=T)
```

```{r}
df <- data.frame(Steps, as.numeric(lev))
```
draw a ggplot
```{r fig.width=9, fig.height=6}
library(ggplot2)
g <- ggplot(df, aes(as.numeric(lev), Steps))
g + geom_line(colour="green")+ggtitle("Time Series Plot of the 5-minute Interval\n and the Average Number of Steps,\n Taken across all Days") + ylab("Average Number of Steps")
```
Imputing missing values
======================
```{r}

missing <- is.na(data)
(n.missing <- sum(missing))

library(mice)
library(Amelia)
set.seed(144)
imputedData <- complete(mice(data))

a.out <- amelia(data, cs="date", m=1)
# Data frame after imputing
 a <- a.out$imputations$imp1
 totStepsImp <- tapply(a$steps, a$date, sum)


# Impute by replacing NA by the mean of steps
data$steps[is.na(data$steps)] <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
# Number of total steps each day
totStepsImp <- tapply(data$steps, data$date, sum)

# Summary statistics before and after imputing
summary(totalSteps)
print(totalSteps)
summary(totStepsImp)
print(totStepsImp)
```

Impute by replacing NA by the mean of steps
```{r}
totstepint<-tapply(data$steps, data$interval, mean, na.rm = TRUE)
print(totstepint)
```
 
 cbind(mean(summary$Steps), mean(summary$Steps))
 
 [1,] 10836 10766
 
 plot the 2 graphs for part one and part 2 histogram
 
```{r fig.width=12, fig.height=6}
par(mfcol=c(1,2))
hist(totstepint, col="red", xlab="Average Total Number of Steps after Imputing", main="", breaks=10)
title("Total Number of steps\n each day After Imputing")
hist(totalSteps, breaks=10, col="green", main="part 1- Total Number of steps each day before imputing", xlab="Average Steps per day")
```

from the data it is clear that missing values have an effect on the total steps. their is a reduction in steps towards the last days


Are there differences in activity patterns between weekdays and weekends?
=========================================================================

```{r}
data$Days <- weekdays(as.Date(as.character(data$date)))
d <- data$Days=="Saturday" | data$Days=="Sunday"
```
Call Saturday and Sunday Weekend
```{r}
data$Days[d] = "Weekend"
```
Call weekdays (Monday through Friday) Weekdays
```{r}
data$Days[!d] = "Weekday" 
activityCopyWD <- data[data$Days == "Weekday",]
activityCopyWE <- data[data$Days == "Weekend",]
```
Convert levels of interval into numeric for weekdays
```{r}
IntervalWD <- as.numeric(levels(as.factor(activityCopyWD$interval)))
IntervalWE <- as.numeric(levels(as.factor(activityCopyWE$interval)))
avgWD <- tapply(activityCopyWD$steps, activityCopyWD$interval, mean)
```
Average number of steps during the weekend
```{r}
avgWE <- tapply(activityCopyWE$steps, activityCopyWE$interval, mean)
dfWD <- data.frame(avgWD, IntervalWD)
dfWE <- data.frame(avgWE, IntervalWE)
data$Days <- as.factor(data$Days)

data$interval <- factor(data$interval)
```

```{r fig.width=10, fig.height=6, echo=T}
# Use base plot
plot(dfWD$IntervalWD, dfWD$avgWD, type='l', main="Comparison of the Average Number of Steps\n between Weekdays and the Weekend", xlab="Interval", ylab="Number of Steps")
lines(dfWE$IntervalWE, dfWE$avgWE, col="red")
legend("topright", c("Weekday", "Weekend"), col=c("yellow", "red"), lty=1)
plot(dfWE$IntervalWE, dfWE$avgWE, type='l', main="Weekend", xlab="Interval", ylab="Number of Steps")
```

using the lattice and drawing the grahps

```{r fig.width=10, fig.height=6, echo=T}

dfWD$wDays <- rep("Weekday", nrow(dfWD))
dfWE$wDays <- rep("Weekend", nrow(dfWD))
# Rename column names to match
colnames(dfWD) <- c("Steps", "Interval", "wDays")
colnames(dfWE) <- c("Steps", "Interval", "wDays")
```
rbind the dataframes
```{r}
df <- rbind(dfWD, dfWE)
```
Convert wDays column into a factor
```{r}
df$wDays <- factor(df$wDays)
library(lattice)
xyplot(Steps ~ Interval | wDays, data=df, type='l', layout=c(1,2), ylab="Average Number of Steps")
```
 Using ggplot2 package
```{r fig.width=10, fig.height=6, echo=T}
g <- ggplot(df, aes(Interval, Steps, fill=wDays, colour=wDays))
g + geom_line() + labs(colour="") + ggtitle("Comparison of the Average Number of Steps\n between Weekdays and Weekend") + ylab("Average Number of Steps")
```


 
 
 
